### 2016-03-17 thesis/introduction.tex

\section{Introduction}
\subsection{Project and Motivation}
Location verification is the process of verifying whether or not a mobile node (computer) is physically present at the location it claims to be. Existing location verification systems attept to provide this service using centralised, trusted ``authoritive'' nodes to provide proof of a mobile node's location. These approaches require investment in infrastructure, and are subject to data breaches and denial of service attacks. This project aims to present a \textit{decentralised} solution to this problem, in which there is no ``authoritive source'' trusted and relied upon to provide and store location information.

This project will focus on obtaining location proofs by using other untrusted mobile nodes as \textit{alibi's}. Proofs will be created over an ad-hoc peer-to-peer network, encrypted, and published on a public append-only bulletin board, known as a \textit{blockchain}. The decentralised nature of the system means that there is no single point of failure, and no one entity controlling the security of every node's location proofs.