### 2016-03-31 presentation.tex
adding initial presentation files

\documentclass{beamer}

\beamertemplatenavigationsymbolsempty
\setbeamertemplate{footline}[frame number]

\usepackage{tikz}
\usetikzlibrary{matrix,positioning,arrows.meta,arrows,calc}

\title[Crisis]
{Decentralised location verification system}
\author{Conor Taylor}
\date{
	B.A.(Mod.) Computer Science\\
	Final Year Project, April 2016\\
	Supervisor: Stephen Barrett
}

\begin{document}

	\frame{\titlepage}

	\begin{frame}
    	\frametitle{Problem}
    	A system that allows participants to verify a users claimed location.
    	\\
    	\null
    	\visible<2->{
    		Goals:
   			\begin{itemize}
    			\item<3-> False location claims must be detectable.
    			\item<4-> Privacy protecting.
    			\item<5-> Cannot rely on any centralised resources.
    			\item<6-> Capable of running in the background on mobile devices.
    		\end{itemize}
    	}
	\end{frame}

	\begin{frame}
		\frametitle{Background}
		There are \textbf{no} known existing decentralised location proof systems.
		\\
		\null
		Existing centralised solutions: hardware and/or software
	\end{frame}

	\begin{frame}
		\frametitle{Design}
		3 distinct entities:
		\begin{itemize}
			\item Mobile node $\vcenter{\hbox{\includegraphics[scale=0.2]{diagrams/overview-presentation/mobile-node.png}}}$
			\item Miner node $\vcenter{\hbox{\includegraphics[scale=0.2]{diagrams/overview-presentation/miner-node.png}}}$
			\item Verifier node $\vcenter{\hbox{\includegraphics[scale=0.2]{diagrams/overview-presentation/verifier-node.png}}}$
		\end{itemize}
	\end{frame}

	\begin{frame}
		\frametitle{Design}
		\framesubtitle{Overview}
		\begin{figure}[H]
			\begin{center}
				\resizebox {\columnwidth} {!} {
					\includegraphics<1>{diagrams/overview-presentation/Overview-13.png}
					\includegraphics<2>{diagrams/overview-presentation/Overview-12.png}
					\includegraphics<3>{diagrams/overview-presentation/Overview-11.png}
					\includegraphics<4>{diagrams/overview-presentation/Overview-10.png}
					\includegraphics<5>{diagrams/overview-presentation/Overview-9.png}
					\includegraphics<6>{diagrams/overview-presentation/Overview-8.png}
					\includegraphics<7>{diagrams/overview-presentation/Overview-7.png}
					\includegraphics<8>{diagrams/overview-presentation/Overview-6.png}
					\includegraphics<9>{diagrams/overview-presentation/Overview-5.png}
					\includegraphics<10>{diagrams/overview-presentation/Overview-4.png}
					\includegraphics<11>{diagrams/overview-presentation/Overview-3.png}
					\includegraphics<12>{diagrams/overview-presentation/Overview-2.png}
					\includegraphics<13>{diagrams/overview-presentation/Overview-1.png}
					\includegraphics<14>{diagrams/overview-presentation/Overview-0.png}
			}
			\end{center}
		\end{figure}
	\end{frame}

	\begin{frame}
		\frametitle{Design}
		\framesubtitle{Identities}
		Used to anonymously identify a node in a transaction.
		\newline

		Balancing goals:
		\begin{itemize}
			\item False location claims must be detectable.
			\item Privacy protecting.
		\end{itemize}
	\end{frame}

	\begin{frame}
		\frametitle{Design}
		\framesubtitle{Identities: Nonce Lists}
		\begin{figure}[H]
			\resizebox {\columnwidth} {!} {\input{diagrams/nonce_list_presentation.tex}}
		\end{figure}
	\end{frame}

	\begin{frame}
		\frametitle{Design}
		\framesubtitle{Identities: Duplication}
		Identity duplication is unavoidable in a decentralised system.
		\begin{table}[]
			\begin{tabular}{l|l}
				\textbf{ID} & \textbf{Contents}\\\hline
				\dots & \\\hline
				ffa0 & \\\hline
				ffa1 & \\\hline
				ffa2 & $T_{A4}$ \visible<2->{, \textcolor{red}{$T_{C102}$}}\\\hline
				ffa3 & \\\hline
				ffa4 & $T_{B87}$\\\hline
				\dots & \\
			\end{tabular}
		\end{table}
	\end{frame}
\end{document}