### 2016-03-31 network_diagram.tex

\documentclass[a4paper,11pt]{article}
\usepackage{amsmath}
\usepackage{tikz}
\usetikzlibrary{matrix,shapes,arrows,positioning,chains, calc}
\begin{document}

\begin{tikzpicture}
\matrix (m)[matrix of nodes, column  sep=2cm,row  sep=8mm, nodes={draw=none, anchor=center,text depth=0pt} ]{
Prover & & Verifier\\[-4mm]
Choose a random number & & \\[-7mm]
$r \in \{0,1\}^{*}.$ & & \\
 & Send $r$ & \\
& & Choose a random number  \\[-7mm]
& &$s \in \{0,1\}^{*}.$\\
 & Send $s$ & \\
};

\draw[shorten <=-1.5cm,shorten >=-1.5cm] (m-1-1.south east)--(m-1-1.south west);
\draw[shorten <=-1.5cm,shorten >=-1.5cm] (m-1-3.south east)--(m-1-3.south west);
\draw[shorten <=-1cm,shorten >=-1cm,-latex] (m-4-2.south west)--(m-4-2.south east);
\draw[shorten <=-1cm,shorten >=-1cm,-latex] (m-7-2.south east)--(m-7-2.south west);
\end{tikzpicture}
\end{document}