### 2016-04-07 thesis/diagrams/distributed_location.tex

\begin{tikzpicture}

\node[draw] (C) at (0,0) {Central node};

\node[draw,style=circle,fill=black] (D1) at (-2,-1) {};
\node[draw,style=circle,fill=black] (D2) at (0,-1) {};
\node[draw,style=circle,fill=black] (D3) at (2,-1) {};

\draw[-] (C) -- (D1) (C) -- (D2) (C) -- (D3);

\node[draw,style=circle] (N1) at (-2.5,-2) {};\node[draw,style=circle] (N2) at (-1.5,-2) {};
\node[draw,style=circle] (N3) at (1.5,-2) {};

\draw[-] (N1) -- (D1) (N2) -- (D1) (N3) -- (D3);

\node[draw,style=circle,fill=black,label=right:Worker node] (L1) at (-1.5,-3) {};
\node[draw,style=circle,label=right:User node] (L2) at (-1.5,-3.5) {};

\end{tikzpicture}