### 2016-04-07 thesis/introduction.tex

\section{Introduction}
\subsection{Project and Motivation}
Location verification is the process of verifying whether or not a  node (computer) is physically present at the location it claims to be. Existing location verification systems attept to provide this service using centralised, trusted ``authoritive'' nodes to provide proof of another node's location. These approaches require investment in infrastructure, and are subject to attacks on privacy and denial of service attacks. This project aims to present a \textit{decentralised} solution to this problem, in which there is no ``authoritive source'' trusted and relied upon to provide and store sensitive location information.

This project describes a decentralised location proof system that operates on mobile devices (mobile nodes). I propose a design in which location proofs are obtained by using other untrusted mobile nodes as \textit{alibi's}. Proofs will be created by communicating over an ad-hoc bluetooth network, encrypted, and published on a public append-only bulletin board, known as a \textit{blockchain}. The decentralised nature of the system means that there is no single point of failure, no entity controlling the security of every node's location proofs, and no entity capable of violating another node's privacy.