### 2016-04-07 thesis/thesis.tex

\title{Decentralised location proof system}

\documentclass[12pt,titlepage]{article}

\usepackage{tikz}
\usetikzlibrary{calc,arrows}
\usepackage{enumitem}
\usepackage{float}

\usepackage{chngcntr}
\counterwithin{figure}{section}

\renewcommand{\thefigure}{\thesection.\arabic{figure}}

\author{Conor Taylor}
\date{
	B.A.(Mod.) Computer Science\\
	Final Year Project, April 2016\\
	Supervisor: Stephen Barrett
}

\begin{document}
\maketitle

\begin{abstract}
\end{abstract}
\newpage

\tableofcontents
\newpage

\listoffigures
\newpage

\input{introduction.tex}
\newpage
\input{background.tex}
\newpage
\input{design.tex}
\newpage
\input{evaluation.tex}

\newpage
\begin{thebibliography}{9}

\bibitem{brassil}
  J. Brassil, P.K. Manadhata,
  ``Verifying the Location of a Mobile Device User'',
  Proc. of MobiSec 2012,
  June 2012.

\bibitem{luo}
  Luo, W., Hengartner, U.,
  ``Proving your Location without giving up your Privacy'',
  Proceedings of the Eleventh Workshop on Mobile Computing Systems \& Applications,
  HotMobile 2010, Annapolis, Maryland, February 22 - 23, pp. 7–12. ACM,
  New York (2010).

\bibitem{khan}
  R. Khan, S. Zawoad, M. M. Haque, and R. Hasan,
  ```Who, When, and Where?' Location Proof Assertion for Mobile Devices'',
  Proceedings of the 28th Annual IFIP WG 11.3 Working Conference on Data and Applications Security and Privacy, ser. DBSec. IFIP,
  July 2014.

\bibitem{otit}
  Khan, R., Zawoad, S., Haque, M., Hasan, R.
  ``OTIT: Towards secure provenance modeling for location proofs'',
  Proc. of ASIACCS. ACM (2014).

\bibitem{sybil}
  Douceur, J.R.,
  ``The sybil attack'',
  Druschel, P., Kaashoek, M.F., Rowstron, A. (eds.) IPTPS 2002. LNCS, vol. 2429, pp. 251–260,
  Springer, Heidelberg (2002).

\bibitem{maxmind}
  MaxMind LLC. GeoIP. http://www.maxmind.com,
  2010.

\bibitem{constraint-based}
  B. Gueye, A. Ziviani, M. Crovella, and S. Fdida,
  ``Constraint-based geolocation of Internet hosts'',
  IEEE/ACM Transactions on Networking,
  14(6):1219–1232, 2006.

\bibitem{location-spoof}
  N. O. Tippenhauer, K. B. Rasmussen, C. Popper, and S. Capkun,
  ``iPhone and iPod location spoofing: Attacks on public WLAN-based positioning systems'',
  SysSec Technical Report,
  ETH Zurich, April, 2008.

\bibitem{tor}
  R. Dingledine, N. Mathewson, and P. Syverson,
  ``TOR: The second generation onion router'',
  Proceedings of the Usenix Security Symposium,
  2004.

\bibitem{distributed}
  Coulouris, George; Jean Dollimore; Tim Kindberg; Gordon Blair,
  ``Distributed Systems: Concepts and Design (5th Edition)'',
  Boston: Addison-Wesley,
  2011.

\bibitem{blueprint}
  M. Swan,
  ``Blockchain: Blueprint for a New Economy'',
  Sebastopol, CA,
  O’Reilly Media, 2015

\bibitem{bitcoin}
  S. Nakamoto,
  ``Bitcoin: A Peer-to-Peer Electronic Cash System'',
  2008

\end{thebibliography}

\end{document}