### 2016-04-18 thesis/design.tex

The location proof system proposed in this project allows \textit{mobile nodes} to create location proofs by initiating an \textit{exchange} with another anonymous mobile node. Mobile nodes, physically moving around, advertise their identity over a short range, ad-hoc network such as bluetooth. When two mobile nodes are in close physical proximity, they initiate an exchange in the background over the ad-hoc network. On completion of the exchange, both nodes can generate their own \textit{transaction}, a cryptographically secure proof of location, attested by the other mobile node (the \textit{alibi}), and make it public.

To create a transaction, two mobile nodes anonymously exchange their GPS coordinates and current timestamp over the ad-hoc network. Both nodes then check that these parameters do not differ from their own observed parameters by more than some value $\epsilon$. Once satisfied, they each create an encrypted, privacy-protecting transaction logging their location and the identity of the the alibi used to create the transaction. This transaction is published onto a public, append-only bulletin board known as a \textit{blockchain} \cite{blueprint}. Once given permission from a mobile node, a \textit{Verifier node} can consult the data in blockchain and determine whether or not the mobile node is present at its claimed location.

Any mobile node acting as an alibi in a transaction will share details of its most recent transactions with the other mobile node in the exchange. This means that when a mobile node seeks verification from a Verifier node, the Verifier can examine the mobile node's alibis, and each of the alibis' recent alibis, forming a large tree of connections which can be used to verify or reject the mobile nodes claimed location (see figure \ref{fig:tree}). Only the owner of the transaction can give permission to a Verifier to check its location history, by providing the Verifier with the keys needed to decrypt the transactions.

Figure \ref{fig:overview} provides an overview of the operation of the proposed system.

\begin{figure}[H]
\begin{center}
\resizebox {\columnwidth} {!} {\includegraphics{diagrams/overview.png}}
\caption{Design and operation overview}
\label{fig:overview}
\end{center}
\end{figure}

\vspace{-1cm}
\begin{enumerate}[label=\textbf{\arabic*}.]
\item Two mobile nodes come in close physical proximity and create an ad-hoc network, over which they initiate an exchange and each create a transaction.
\item Both mobile nodes publish their own transaction to a Miner node.
\item The Miner node who received the transaction distributes it to the other miner nodes in the network.
\item One Miner node solves the proof-of-work for a block of transactions, and the new block is appended to the blockchain.
\item A mobile node requests verification from a Verifier node (request structure is explained in section \ref{ssec:verification}).
\item The Verifier node refers to the public blockchain to examine the requesting node's transactions and determine their validity.
\item The Verifier accepts or rejects the  mobile node's claimed location based on its analysis of the transactions in the blockchain. 
\end{enumerate}

\section{Identities}
In order to preserve users' privacy, a user will only ever use an identity for one transaction, before generating a new one. This prevents curious users from watching the public blockchain for a known identity and tracking it. However, to prevent identity theft, it is important for a Verifier to be able to prove that a mobile node is the owner of each identity. This is achieved using a public/private key pair for each mobile node.

When a mobile node first joins the network, it generates a public/private key pair. To maintain anonymity, the mobile node will use the public key to encrypt \textit{nonces} to create \textit{identities}, and use these to identify itself in a transaction. A different identity will be created for each transaction. During the verification stage, a node will provide a Verifier with its public key, along with the list of the nonces used to generate its $n$ most recent transactions (see figure \ref{fig:verify_request}). This allows a Verifier to calculate all of the node's identities, and prove that they were all created by the same node. A Verifier can then retrieve the transactions associated with those identities from the blockchain. Verifiers will also ensure that a node owns the private key associated with its provided public key, to prevent a weak identity attack.

\subsection{Nonce list} \label{sssec:nonce_list}
A \textit{Nonce List} is simply a list of all of the nonces a mobile node has used so far to generate identities. When a mobile node wishes to generate a new identity, it must first generate a new random nonce which is then encrypted to create the identity. The nonce is appended to the mobile node's nonce list, so it can be used later during verification (see section \ref{ssec:verification}).

\begin{figure}[H]
\resizebox {\columnwidth} {!} {\input{diagrams/nonce_list.tex}}
\caption{Identity generation from Nonce List}
\label{fig:nonce_list}
\end{figure}

In figure \ref{fig:nonce_list} above, nonce $0047$ is encrypted with $K^+$, a node's public key, to create identity $ee3bc14$. By providing the nonce list and $K^+$, a node can prove that it created all of the identities in the identities list. It must also verify that it owns $K^-$, its private key.

\subsection{Identity duplication}
Since all identities are generated from random nonces, there is nothing to prevent two nodes from generating the same identity and using it in different transactions. It is therefore possible that during verification, a Verifier node may find multiple transactions with the same ID while searching the blockchain. Verifiers are provided with decryption key from the requesting node for the transactions they are searching for, which will only successfully decrypt one of the transactions with the duplicate ID. A Verifier will try to decrypt each one until it finds the one that can be decrypted successfully.

\newpage
\section{Transactions} \label{sec:transactions}
During an exchange between two nodes, a number of different items must be shared and calculated. Figure \ref{fig:transaction} describes and explains an honest, successful exchange, assuming an ad-hoc network has already been set up between nodes $A$ and $B$.

\begin{figure}[H]
\begin{center}
\resizebox {1.1\columnwidth} {!} {\input{diagrams/transaction.tex}}
\caption{Honest, successful exchange}
\label{fig:transaction}
\end{center}
\end{figure}

\begin{itemize}
	\item[] \textbf{(1)} Before the exchange begins, both nodes generate a new entry in their nonce list (e.g. $NL_A[n]$), and generate a new identity from that nonce (e.g. $ID_A[n]$).
	\item[] \textbf{(2)} Node $A$ sends its identity to node $B$, along with its observed current timestamp $ts_A$ and current GPS location $loc_A$.
	\item[] \textbf{(3)} Node $B$ verifies that the observed timestamp and GPS location received from node $A$ are within an acceptable difference from its own observations. This difference is defined by some system constant $\epsilon$.
	\item[] \textbf{(4)} If node $B$ verifies the parameters received from node $A$, it will respond with its own identity and observations for node $A$ to verify.
	\item[] \textbf{(5)} Node $A$ verifies that the parameters received from node $B$ differ by no more than $\epsilon$ from its own observed parameters.
	\item[] \textbf{(6)} Node $A$ sends $KP_{An}$, its $nth$ \textit{Key Packet}. Key Packets are used to provide alibi credibility, and are discussed in section \ref{sssec:key_packets}.
	\item[] \textbf{(7)} Node $B$ responds with its own Key Packet.
	\item[] \textbf{(8)} Both nodes generate their own transaction $T$, using their alibi's Key Packets. These are then used to generate $P$, the data published onto the public blockchain.
\end{itemize}

\subsection{Key Lists} \label{ssec:key_lists}
When a node is initialised, it must begin to record two \textit{Key Lists} for itself. A \textit{Key List} is a list of encryption keys previously used for transactions. For example, $KL_{AT}[n]$ and $KL_{AL}[n]$ are the two keys used to encrypt node $A$'s $n$th transaction and matching chronological link, respectively. These \textit{Key Lists} are assumed to have been initialised and populated before the transaction described in Figure \ref{fig:transaction} begins.

\subsection{Transaction creation}
Once nodes have exchanged \textit{Key Packets}, their communication is finished and they can close their ad-hoc channel. Each node creates a new transaction, $T$, and generates data $P$ to publish onto the public blockchain. The following transaction will be created by node $A$, after communicating with node $B$:
\\

$T_{An} = KL_{AT}[n](ts|loc|ID_{A}[n]|ID_{B}[m]|KP_{Bm})$
\\

Where:
\begin{itemize}[noitemsep,topsep=0pt]
	\item[] $\mathbf{n}$ is node $A$'s current transaction number (i.e. node $A$ has published $n-1$ transactions before now).
	\item[] $\mathbf{KL_{AT}}$ is node $A$'s transaction \textit{Key List}, used to encrypt transactions. $KL_{AT}[n]$ is the $nth$ entry in the Key List.
	\item[] $\mathbf{ts}$ is the timestamp of the transaction, as recorded by node $A$.
	\item[] $\mathbf{loc}$ is the GPS coordinates of the transaction, as recorded by node $A$.
	\item[] $\mathbf{ID_{A}[n]}$ is $A$'s $nth$ identity, used for this transaction only.
	\item[] $\mathbf{ID_{B}[m]}$ is $B$'s $mth$ identity.
	\item[] $\mathbf{KP_{Bm}}$ is $B$'s \textit{Key Packet} at time $m$. Its exact contents are discussed in more detail in section \ref{sssec:key_packets}
\end{itemize}

\subsection{Publishing to the blockchain}
After creating $T_{An}$, node $A$ must publish it to the public blockchain. The data published must be searchable in order to provide a Verifier with some way of identifying specific transactions. Nodes use their current ID as the identifier for a transaction.

\textit{Backwards-chaining} is an important part of verifiability. A Verifier must be able to prove that the proof chain provided by a mobile node has not been chronologically re-ordered. This was designed to satisfy the ``order preserving'' property of OTIT \cite{otit}. Each transaction must therefore be accompanied by a link to the node's chronologically previous transaction. This link is encrypted with $KL_{AL}[n]$, node $A$'s $nth$ \textit{link} key list, to prevent a curious user from monitoring the public blockchain to observe a user's transaction publishing activity. The timestamp $ts_A$ is also included in the backwards-chaining data, required to verify the ``order preserving'' property of OTIT. Node $A$'s transaction data $P_{An}$ will therefore be in the form:
\\

$P_{An} = ID_{An}|KL_{AL}[n](ID_{A}[n-1]|ts_A)|T_{An}$

\null
To publish $P_{An}$, node $A$ must send $P_{An}$ to a miner node, who will add it to its \textit{pending transaction} queue, then send it to other miners in the network to add it to their queues. The transaction will be signed into the next block in the blockchain, once the next proof-of-work is solved.

Node $B$, $A$'s alibi for the above transaction example, will publish $P_{Bm}$ after the exchange. If node $B$'s transaction is not available on the blockchain, node $A$'s transaction will be considered invalid and disregarded during verification. A transaction $T_A$ in the blockchain must be accompanied by a reference to some valid alibi transaction $T_B$, and $T_B$ must reference $T_A$, otherwise the proof is not considered credible by a Verifier.

\subsection{Key Packets} \label{sssec:key_packets}
A key packet is more than simply a list of keys that a mobile node has used to encrypt its transactions; in order to preserve the ``selective in-sequence privacy'' property of OTIT \cite{otit}, a user must be able to choose transactions that it does not want others to be able to decrypt. However, in order to simultaneously preserve OTIT's ``privacy protected chronology'' property, the transaction backwards-chaining must not be broken. This is why two separate \textit{Key Lists} are maintained; $KL_{AL}$ is used to encrypt the backwards-chaining link and transaction timestamp, while $KL_{AT}$ encrypts the transaction data. For example, if node $B$ does not want to reveal transaction $T_{Bm-1}$, it would use the following key packet:
\\

${KP_{Bm} = (\{KL_{BL}[m], KL_{BT}[m]\}, KL_{BL}[m-1], \{KL_{BL}[m-2], KL_{BT}[m-2]\})}$

\null
In this case, $KL_{BT}[m-1]$ has not been included with $KL_{BL}[m-1]$, in order to protect the privacy of this transaction. This means that a node who receives this key packet will not be able to decrypt $B$'s transaction $T_{Bm-1}$, but can still prove that a transaction exists at that time and in the correct sequence, as $B$ will still provide $KL_{BL}[m-1]$ (see figure \ref{fig:link}). This may be needed, for example, if a user does not want to allow anyone to decrypt any transactions it has created within 5km of its home, in order to protect its privacy.

\begin{figure}[H]
\begin{center}
\resizebox {\columnwidth} {!} {\input{diagrams/link.tex}}
\caption{A proof chain with transaction $T_{m-1}$ hidden}
\label{fig:link}
\end{center}
\end{figure}

\subsection{Aborting exchanges} \label{ssec:aborting_exchanges}
A node may decide to abort an exchange if it thinks it is in contact with a malicious node. In figure \ref{fig:aborted_transaction}, node $A$ encounters a malicious node $M$, who tries to spoof its location to node $A$. Node $A$ will abort the exchange once it notices that $|loc_A-loc_M| \geq \epsilon_{loc}$.

To abort the exchange, node $A$ terminates the ad-hoc network with node $M$, and won't publish anything onto the blockchain. This means that even if node $M$ fabricates some data from node $A$ and publishes a transaction onto the blockchain, the matching transaction from node $A$ will not be present, so node $M$'s transaction will be disregarded by any Verifiers.

\begin{figure}
\resizebox {\columnwidth} {!} {\input{diagrams/aborted_transaction.tex}}
\caption{An aborted transaction due to malicious node $M$}
\label{fig:aborted_transaction}
\end{figure}

\clearpage
\section{Verification} \label{ssec:verification}
In the verification stage, a mobile node attempts to prove its location to a Verifier node, e.g. a bank. To do this, the mobile node sends the Verifier a number of parameters, as shown in Figure \ref{fig:verify_request}.

\begin{figure}[H]
\begin{center}
\resizebox {0.8\columnwidth} {!} {\input{diagrams/verify_request.tex}}
\caption{Verification request}
\label{fig:verify_request}
\end{center}
\end{figure}

\begin{itemize}[noitemsep,topsep=0pt]
	\item[] $\mathbf{ID_{A}[n-1]}$ is the ID of node $A$'s most recently published transaction. Note that $ID_A[n]$ has already been generated, but hasn't been used in a transaction yet. Node $A$ has generated it so that it is ready to be used in the next transaction.
	\item[] $\mathbf{KP_{An-1}}$ is node $A$'s key packet up to transaction $n-1$, and includes keys for as many transactions as node $A$ feels is appropriate in order to receive verification.
	\item[] $\mathbf{loc}$ is node $A$'s current claimed GPS coordinates for which it is seeking verification.
	\item[] $\mathbf{NP_{An-1}}$ is node $A$'s \textit{Nonce Packet}, a subset of its \textit{Nonce List} (section \ref{sssec:nonce_list}). A Nonce Packet is the list of the nonces used to generate node $A$'s most recent IDs, up to $n-1$, that node $A$ is willing to share with the Verifier. 
\end{itemize}

\null
The goal of the verification stage is for the Verifier to either accept or reject the location that the user is claiming. Given $ID_{A}[n-1]$ and $KP_{An-1}$, a Verifier is able to retrieve node $A$'s previous transactions, and the transactions of all of its alibis. This will allow the Verifier to recursively inspect the transactions of all alibis used in a node's transactions, until it is satisfied that the location can be verified or rejected. This recursive inspection can be visualised as a tree, as shown in figure \ref{fig:tree}.

\begin{figure}[H]
\begin{center}
\resizebox {0.7\columnwidth} {!} {\input{diagrams/tree.tex}}
\vspace{-3cm}
\caption{Alibi chain relationship tree}
\label{fig:tree}
\end{center}
\end{figure}

If the verifier wishes to recurse more, it could inspect the black node's alibis' transaction chains. Note that the example in figure \ref{fig:tree} is a simple one, with $n = 3$ being the length of the transaction chain subset provided for verification. In reality $n$ should be much larger.

There are a huge number of possible factors involved in reaching a conclusion on the verifiability of a user from the data in the blockchain; some of these factors will likely be unique to certain Verifiers, and kept secret to improve reliability and avoid gaming by malicious users. Some simple example factors include:
\begin{itemize}
	\item \textbf{Alibi credibility:} if each of the node’s alibis have little or no alibis themselves, they will not be considered credible alibis, and a Verifier will reject the verification request (see figure \ref{fig:uncredible_tree}).
	\item \textbf{Transaction relevance:} if a Verifier considers the transactions it receives to be out of date, the request will be rejected on the grounds that the transactions are no longer relevant.
	\item \textbf{Alibi reuse:} if a Verifier can prove that alibis are being reused frequently, then it will reject the verification request. Due to transaction anonymity, a Verifier may not always be able to tell if two alibis are the same person or not. e.g. node $A$ could use node $B$ as an alibi, then wait for node $B$ to complete $n$ more transactions, then use node $B$ again.
	This is not a viable attack on the verification system. Depending on the size of $n$, enough time may pass between both uses of node $B$ as an alibi that the transactions are no longer relevant.
\end{itemize}