### 2016-04-23 thesis/thesis.tex

\documentclass[12pt]{report}

\title{Decentralised location proof system}

\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{setspace}

\usepackage{tikz}
\usetikzlibrary{matrix,positioning,arrows.meta,arrows,calc}
\usepackage{enumitem}
\usepackage{float}
\usepackage{subfig}

\author{Conor Taylor}
\date{
	B.A.(Mod.) Computer Science\\
	Final Year Project, April 2016\\
	Supervisor: Stephen Barrett
}

\begin{document}

\onehalfspacing

\pagenumbering{roman}
\maketitle

\renewcommand{\abstractname}{Abstract}
\begin{abstract}
Location proof systems are systems which allow a user to undeniably prove its location to another user. Existing systems have thus far been designed in a centralised manner, using a trusted central server to provide honest location proofs to users. This presents privacy and reliability issues, as these systems rely on the honesty and availability of the central server. This report presents the design and evaluation of a decentralised location proof system, which overcomes the shortcomings of existing centralised location proof systems.
\end{abstract}

\renewcommand{\abstractname}{Acknowledgements}
\begin{abstract}
I wish to thank my supervisor Stephen Barrett, for his continuous guidance and enthusiasm throughout this project. I would also like to thank Michael Clear for his excellent advice and suggestions, and Douglas Temple for his generous support.
\end{abstract}

\tableofcontents
\newpage

\listoffigures
\newpage

\pagenumbering{arabic}
\chapter{Introduction} \label{ch:introduction}
\input{introduction.tex}

\chapter{Background} \label{ch:background}
\input{background.tex}

\chapter{Design} \label{ch:design}
\input{design.tex}

\chapter{Evaluation} \label{ch:evaluation}
\input{evaluation.tex}

\chapter{Conclusion and Future Work} \label{ch:conclusion}
\input{conclusion.tex}

\newpage
\begin{thebibliography}{9}

\bibitem{brassil}
  J. Brassil, P.K. Manadhata,
  ``Verifying the Location of a Mobile Device User'',
  Proc. of MobiSec 2012,
  June 2012.

\bibitem{luo}
  W. Luo, U. Hengartner,
  ``Proving your Location without giving up your Privacy'',
  Proceedings of the Eleventh Workshop on Mobile Computing Systems \& Applications,
  HotMobile 2010, Annapolis, Maryland, February 22 - 23, pp. 7–12. ACM,
  New York (2010).

\bibitem{khan}
  R. Khan, S. Zawoad, M. Haque, and R. Hasan,
  ```Who, When, and Where?' Location Proof Assertion for Mobile Devices'',
  Proceedings of the 28th Annual IFIP WG 11.3 Working Conference on Data and Applications Security and Privacy, ser. DBSec. IFIP,
  July 2014.

\bibitem{otit}
  R. Khan, S. Zawoad, M. Haque, and R, Hasan.
  ``OTIT: Towards secure provenance modeling for location proofs'',
  Proc. of ASIACCS. ACM (2014).

\bibitem{sybil}
  J. Douceur,
  ``The Sybil Attack'',
  In First IPTPS,
  March 2002.

\bibitem{maxmind}
  MaxMind LLC. GeoIP. http://www.maxmind.com,
  2010.

\bibitem{constraint-based}
  B. Gueye, A. Ziviani, M. Crovella, and S. Fdida,
  ``Constraint-based geolocation of Internet hosts'',
  IEEE/ACM Transactions on Networking,
  14(6):1219–1232, 2006.

\bibitem{location-spoof}
  N. O. Tippenhauer, K. B. Rasmussen, C. Popper, and S. Capkun,
  ``iPhone and iPod location spoofing: Attacks on public WLAN-based positioning systems'',
  SysSec Technical Report,
  ETH Zurich, April, 2008.

\bibitem{tor}
  R. Dingledine, N. Mathewson, and P. Syverson,
  ``TOR: The second generation onion router'',
  Proceedings of the Usenix Security Symposium,
  2004.

\bibitem{distributed}
  G. Coulouris, J. Dollimore, T. Kindberg, and G. Blair,
  ``Distributed Systems: Concepts and Design (5th Edition)'',
  Boston: Addison-Wesley,
  2011.

\bibitem{blueprint}
  M. Swan,
  ``Blockchain: Blueprint for a New Economy'',
  Sebastopol, CA,
  O’Reilly Media, 2015

\bibitem{bitcoin}
  S. Nakamoto,
  ``Bitcoin: A Peer-to-Peer Electronic Cash System'',
  2008

\bibitem{femtocell}
  V. Chandrasekhar, J. Andrews, and A. Gatherer,
  ``Femtocell Networks: A Survey'',
  IEEE Communications Magazine, Vol. 46, No. 9, pps. 59-67,
  September 2008

\bibitem{ddos}
  J. Mirkovic, P. Reiher,
  ``A taxonomy of DDoS attack and DDoS defense mechanisms'',
  ACM SIGCOMM Comput. Commun. Rev. 34, 2, 39–53,
  2004

\bibitem{weboftrust}
  R. V. Guha, R. Kumar, P. Raghavan, and A. Tomkins,
  ``Propagation of trust and distrust'',
  in Proceedings of WWW, pp. 403–412,
  2004.

\end{thebibliography}

\end{document}